Proving Atomic Specifications for Data Structures with Fine-Grained Locking\thanks{Supported by the NSF (details).}

### 2022-04-07 concurrency/ASL 2022/intro.tex
folder for ASL submission

% Contributions:
% verified C implementation of BST with atomic specs -- the first? VeriFast did it, but not atomic specs
% technique for connecting locks with abstract state/atomicity
% VST-style decomposition of proofs into C-to-local-abstract and local-to-global-abstract
% delete proofs -- interesting interplay between non-visible range increase and lock-protected range modification
% lock-free implementation?

@@

Concurrent separation logic~\cite{csl} is a useful tool for proving correctness of concurrent programs. While early iterations focused on data structures implemented with locks~\cite{gotsman, oraclesematic} and specifications involving memory safety and data-race-freedom, more recent logics are able to prove linearizability-style functional correctness~\cite{tada} of lock-free implementations that use low-level atomic memory operations~\cite{iris}, even under weak memory models~\cite{gps, rsl}. In particular, \emph{logical atomicity}~\cite{tada} has become the gold standard for concurrent data structure specifications, capturing the idea that data structure operations should appear to take place instantaneously with no externally visible intermediate states. The effects of a logically atomic operation become visible at a \emph{linearization point}, a single instruction (usually an atomic memory operation or call to a logically atomic function) that publishes the new state of the data structure.

>>>>

% Contributions:
% verified C implementation of BST with atomic specs -- the first? VeriFast did it, but not atomic specs
% technique for connecting locks with abstract state/atomicity
	% note that the flow interfaces team sort of does atomicity for locks, but they assume an atomic spec for locks that's also application-specific? no, they implement them in the TaDA style and prove an atomic spec.
% VST-style decomposition of proofs into C-to-local-abstract and local-to-global-abstract -- this corresponds to flow interfaces/CSS templates, basically
% delete proofs -- interesting interplay between non-visible range increase and lock-protected range modification
% lock-free implementation?

@@

Concurrent separation logic~\cite{csl} is a useful tool for proving correctness of concurrent programs. While early iterations focused on data structures implemented with locks~\cite{gotsman,oraclesematic} and specifications involving memory safety and data-race-freedom, more recent logics are able to prove linearizability-style functional correctness~\cite{tada} of lock-free implementations that use low-level atomic memory operations~\cite{iris}, even under weak memory models~\cite{gps}%update cite
. In particular, \emph{logical atomicity}~\cite{tada} has become the gold standard for concurrent data structure specifications, capturing the idea that data structure operations should appear to take place instantaneously with no externally visible intermediate states. The effects of a logically atomic operation become visible at a \emph{linearization point}, a single instruction (usually an atomic memory operation or call to a logically atomic function) that publishes the new state of the data structure.
